---
title: "Data assembly"
author: "Carole Voulgaris"
date: "1/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(sf)
library(fs)
library(lubridate)
library(tidycensus)
library(here)
library(arrow)
```


# Load Smart Location Database

Downloaded from https://edg.epa.gov/EPADataCommons/public/OA/SLD/SmartLocationDatabaseV3.zip and saved to raw_datasets folder (listed in .gitignore)

```{r}
sld_data <- st_read(path('raw_datasets',
                            'SmartLocationDatabaseV3',
                            'SmartLocationDatabase.gdb')) %>%
  filter(STATEFP == '42',
         COUNTYFP == '003') %>%
  select(D1A, # HU density
         D1C8_RET, # retail job density
         D1C8_OFF, # office job density
         D1C8_IND, # industrial
         D1C8_SVC, # service
         D1C8_ENT, # entertainment
         D1C8_ED, # education
         D1C8_HLTH, # health
         D1C8_PUB, # government?
         D3APO, # ped network density
         D4A, # distance to transit
         D4D, # transit frequency per square mile
         D5AR, # jobs within 45 minutes, car
         D5BR) %>% # jobs within 45 minutes, transit
  rename(hu_dens = D1A,
         retail_dens = D1C8_RET,
         office_dens = D1C8_OFF,
         ind_dens = D1C8_IND,
         svc_dens = D1C8_SVC,
         entrtnmnt_dens = D1C8_ENT,
         educ_dens = D1C8_ED,
         hlth_dens = D1C8_HLTH,
         pub_dens = D1C8_PUB,
         ped_network = D3APO,
         transit_dist = D4A,
         transit_freq = D4D,
         car_jobs = D5AR,
         transit_jobs = D5BR) %>%
  st_transform("WGS84")            
```

# Load county assessor data

```{r}
parcel_data <- read_csv('https://data.wprdc.org/dataset/2b3df818-601e-4f06-b150-643557229491/resource/f2b8d575-e256-4718-94ad-1e12239ddb92/download/assessments.csv') %>%
  filter(CLASSDESC == "RESIDENTIAL") 
 

parcels <- st_read('https://data.wprdc.org/dataset/6bb2a968-761d-48cf-ac5b-c1fc80b4fe6a/resource/42231cab-8341-48d6-b695-47612dd6514a/download/parcelcoords.csv',
  options = c("X_POSSIBLE_NAMES=x", "Y_POSSIBLE_NAMES=y")) %>%
  rename(PARID = PIN) %>%
  inner_join(parcel_data) %>%
  st_set_crs("WGS84")

parcel_info <- parcels %>%
  select(PROPERTYHOUSENUM, PROPERTYFRACTION, PROPERTYADDRESS,
         PROPERTYUNIT, PROPERTYCITY, PROPERTYSTATE, PROPERTYZIP)
```

Selected parcel variables --

Exclude properties that have no data on the most recent sale. 
Exclude properties where the most recent sale is before the second-most-recent sale.

For purposes of calculating the time between sales:
* If there is no data on the second-most recent sale, assume if happened on January 1, 1806 (the earliest sale in the dataset is November 24, 1806) for $100
* Average time between sales over two periods only if there is data on the third most recent sale (PREVSALEDATE2)


```{r}
parcel_values <- parcels %>%
  filter(!is.na(SALEDATE)) %>%
  mutate(PREVSALEDATE = ifelse(is.na(PREVSALEDATE), 
                               "01-01-1806", 
                               PREVSALEDATE)) %>%
  mutate(SALEDATE = mdy(SALEDATE),
         PREVSALEDATE = mdy(PREVSALEDATE),
         PREVSALEDATE2 = mdy(PREVSALEDATE2),
         ASOFDATE = dmy(ASOFDATE)) %>%
  mutate(years_since_sale = as.numeric(today() - SALEDATE)/365.25,
         btw_sales_1 = as.numeric(SALEDATE - PREVSALEDATE)/365.25,
         btw_sales_2 = as.numeric(PREVSALEDATE - PREVSALEDATE2)/365.25) %>%
  filter(btw_sales_1 > 0) %>%
  mutate(btw_sales_avg = ifelse(is.na(btw_sales_2), 
                                btw_sales_1,
                                (btw_sales_1 + btw_sales_2)/2)) %>%
  mutate(ten_year_date = case_when(PREVSALEDATE == mdy("01-01-1806") ~ SALEDATE,
                                   SALEDATE < (today() - years(10)) ~ SALEDATE,
                                   PREVSALEDATE < (today() - years(10)) ~ PREVSALEDATE,
                                   !is.na(PREVSALEPRICE2) ~ PREVSALEDATE2,
                                   TRUE ~ PREVSALEDATE)) %>%
  mutate(ten_year_price = case_when(PREVSALEDATE == mdy("01-01-1806") ~ SALEPRICE,
                                    SALEDATE < (today() - years(10)) ~ SALEPRICE,
                                    PREVSALEDATE < (today() - years(10)) ~ PREVSALEPRICE,
                                    !is.na(PREVSALEPRICE2) ~ PREVSALEPRICE2,
                                    TRUE ~ PREVSALEPRICE)) %>%
  filter(!is.na(ten_year_price)) %>%
  mutate(ten_year_period = as.numeric(ASOFDATE - ten_year_date)/365.25) %>%
  mutate(ten_year_price = ifelse(ten_year_price < 100, 100, ten_year_price)) %>%
  mutate(ten_yr_increase_rate = ((FAIRMARKETTOTAL/ten_year_price)^(1/ten_year_period)) - 1) %>%
  mutate(bldg_pct = FAIRMARKETBUILDING / FAIRMARKETTOTAL) %>%
  select(PARID,
         FAIRMARKETTOTAL, 
         bldg_pct, 
         btw_sales_avg,
         ten_yr_increase_rate)
```

join parcel data to SLD data 

```{r}
data_pts1 <- st_join(parcel_values, sld_data) %>%
  filter(!is.na(hu_dens))
```

Get thresholds for building age

```{r}
vars <- c(
  since_2014 = "B25034_002",
  b_2010_to_2013 = "B25034_003",
  b_2000_to_2009 = "B25034_004",
  b_1990_to_1999 = "B25034_005",
  b_1980_to_1989 = "B25034_006",
  b_1970_to_1979 = "B25034_007",
  b_1960_to_1969 = "B25034_008",
  b_1950_to_1959 = "B25034_009",
  b_1940_to_1949 = "B25034_010",
  before_1940 = "B25034_011"
)

chk_bldg_age <- get_acs(geography = "county",
                    state = "PA",
                    variables = vars,
                    summary_var = "B25034_001") %>%
  filter(GEOID == "42003") %>%
  mutate(pct = estimate / summary_est) %>%
  mutate(cum_pct = cumsum(pct))
```


Get census data

* percent renters
* percent owners without mortgage
* monthly ownership cost with mortgage (median)
* monthly ownership cost without mortgage (median)
* monthly rent (median)
* income (gini)
* pct white
* diversity index (race)
* diveristy index (building age)

```{r}
vars <- c(total_hus = "B25003_001",
          renter_occ = "B25003_003",
          no_mort = "B25081_008",
          med_cost_with_mort = "B25088_002",
          med_cost_no_mort = "B25088_003",
          median_rent = "B25064_001",
          gini = "B19083_001",
          total_pop = "B03002_001",
          nh_white = "B03002_003",
          nh_black = "B03002_004",
          nh_asian = "B03002_006",
          hispanic = "B03002_012",
          since_2014 = "B25034_002",
          b_2010_to_2013 = "B25034_003",
          b_2000_to_2009 = "B25034_004",
          b_1990_to_1999 = "B25034_005",
          b_1980_to_1989 = "B25034_006",
          b_1970_to_1979 = "B25034_007",
          b_1960_to_1969 = "B25034_008",
          b_1950_to_1959 = "B25034_009",
          b_1940_to_1949 = "B25034_010",
          before_1940 = "B25034_011")

census <- get_acs(geography = "tract", 
                  variables = vars, 
                  state = "PA",
                  county = "Allegheny",
                  output = "wide", 
                  geometry = TRUE) %>%
  mutate(pct_rental = renter_occE / total_husE,
         pct_no_mort = no_mortE / total_husE,
         pct_nh_white = nh_whiteE / total_popE,
         pct_nh_black = nh_blackE / total_popE,
         pct_nh_asian = nh_asianE / total_popE,
         pct_hispanic = hispanicE / total_popE,
         pct_since_1980 = (since_2014E + 
                          b_2010_to_2013E +
                          b_2000_to_2009E +
                          b_1990_to_1999E +
                          b_1980_to_1989E) / total_husE,
         pct_60s_70s = (b_1970_to_1979E +
                          b_1960_to_1969E) / total_husE,
         pct_40s_50s = (b_1950_to_1959E +
                          b_1940_to_1949E) / total_husE,
         pct_pre_1940 = before_1940E / total_husE) %>% 
  mutate(h_white = pct_nh_white * log(pct_nh_white),
         h_black = pct_nh_black * log(pct_nh_black),
         h_asian = pct_nh_asian * log(pct_nh_asian),
         h_hisp = pct_hispanic * log(pct_hispanic),
         h_since_1980 = pct_since_1980 * log(pct_since_1980),
         h_60s_70s = pct_60s_70s * log(pct_60s_70s),
         h_40s_50s = pct_40s_50s * log(pct_40s_50s),
         h_pre_40s = pct_pre_1940 * log(pct_pre_1940)) %>%
  rowwise(GEOID) %>%
  mutate(race_entropy = -1*sum(c(h_white, h_black, h_asian, h_hisp), na.rm = TRUE),
         bldg_entropy = -1*sum(c(h_since_1980, 
                                 h_60s_70s, 
                                 h_40s_50s,
                                 h_pre_40s), na.rm = TRUE)) %>%
  ungroup() %>%
  rename(med_cost_with_mort = med_cost_with_mortE,
         med_cost_no_mort = med_cost_no_mortE,
         median_rent = median_rentE,
         gini = giniE) %>%
  select(GEOID, pct_rental, pct_no_mort, med_cost_with_mort, 
         med_cost_no_mort, median_rent, gini, pct_nh_white,
         race_entropy, bldg_entropy) %>%
  st_transform("WGS84")
```



```{r}
# filter if values are missing for less than three percent of all parcels
data_pts2 <- st_join(data_pts1, census) 

data_pts2 <- data_pts2 %>%
  filter(!is.na(data_pts2$median_rent) &
           !is.na(data_pts2$med_cost_no_mort) &
           !is.na(data_pts2$med_cost_with_mort) &
           !is.na(data_pts2$pct_no_mort) &
           !is.na(data_pts2$pct_rental))
```

write to file

```{r}
full_data_df <- st_drop_geometry(data_pts2)
parcels <- data_pts2 %>%
  select(PARID)

write_parquet(full_data_df, here("assembled-data",
                                 "easy-data.parquet"))
st_write(parcels, here("assembled-data",
                       "parcel-locs.geojson"))
```
