---
title: "Data assembly"
author: "Carole Voulgaris"
date: "1/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, results='hide'}
options(java.parameters = '-Xmx2G')

library(tidyverse)
library(sf)
library(fs)
library(lubridate)
library(psych)
library(nFactors)
library(tigris)
library(r5r)
library(here)
library(arrow)
library(GPArotation)
library(ggfortify)
```



Check if factor analysis would be okay

```{r}
full_data <- read_parquet("assembled-data/easy-data.parquet") %>%
  inner_join(read_parquet("assembled-data/parcel-access.parquet")) %>%
  inner_join(read_parquet("assembled-data/loan-data.parquet")) 

factor_data <- full_data %>%
  select(-PARID, -GEOID, -id)

drop_na(full_data)
  
KMO(factor_data)
```
KMO is greater than 0.6, so factor analysis is okay

```{r}
cortest.bartlett(factor_data)
```

p-value is zero, so factors analysis is okay

Take a guess at the right number of factors

```{r}
# PCR calculates correlation matrix first and finds eigenvector and eigenvalue based on the correlation matrix.
# conduct PCR analysis with the correlation matrix to reduce calculation time.
correlation = cor(factor_data, use = "complete.obs" )
corPlot(correlation)

# PC number is 11, Factor component number is 4
fa.parallel(correlation)
abline( h = 0 )

# VSS(correlation)
# parallel <- fa.parallel(factor_data)
```

Do principal component analysis and factor analysis with different numbers of factors (as low as 5, as high as the recommended number) and each possible rotation:

* "none"
* "varimax"
* "quartimax"
* "bentlerT" not worked
* "equamax"
* "varimin"
* "geominT"
* "bifactor"
* "Promax"
* "promax"
* "oblimin"
* "simplimax"
* "bentlerQ" not worked
* "geominQ"
* "biquartimin"
* "cluster" not worked

Identify the best-fitting 

```{r}
# conduct PCR analysis using different nfactors and rotation methods
for (i in c(1:15)) {
  filename <- paste0("pca_", i)
  assign(
    filename,
    pca(correlation,
        nfactors = i,
        rotate = "none"))
}

for (i in c("none", "varimax", "quartimax",  "equamax", "varimin", "geominT", "bifactor", "Promax", "promax", "oblimin", "simplimax",  "geominQ", "biquartimin")) {
  filename <- paste0("pca_", i)
  assign(
    filename,
    pca(correlation,
        nfactors = 11,
        rotate = i)
  )
}

# fit improves as the nfactors increases (it will be 1 if I include 39 variables). Also, rotations do not change the fit  
pca_1$fit
pca_2$fit
pca_3$fit
pca_4$fit
pca_5$fit

pca_none$fit
pca_varimax$fit
pca_quartimax$fit
pca_equamax$fit
pca_geominT$fit
```

```{r}
# extract loads (eigenvectors)

pca_load <- pca_none$loadings[, 1:11]

ggplot(pca_load, aes(PC1, PC2)) +
  geom_point() +
  geom_text(aes(label=ifelse(PC1>0.5 | PC1 < -0.5, row.names(pca_load), ""),hjust=0,vjust=0))
```

```{r}
ggplot(pca_load) +
  geom_bar(aes(reorder(row.names(pca_load), PC1), PC1), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# PC1: commercial area (city center) vs single-family residential area
# transit_jobs and transit_dist seem containing some errors (-9999)
```


```{r}
ggplot(pca_load) +
  geom_bar(aes(reorder(row.names(pca_load), PC2), PC2), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# PC2: wealth
```


```{r}
ggplot(pca_load) +
  geom_bar(aes(reorder(row.names(pca_load), PC3), PC3), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# PC3: CBD
```


```{r}
ggplot(pca_load) +
  geom_bar(aes(reorder(row.names(pca_load), PC4), PC4), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# PC4: diversity and housing access
```


```{r}
ggplot(pca_load) +
  geom_bar(aes(reorder(row.names(pca_load), PC5), PC5), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# PC5: new residential area vs old residential area
```

```{r}
# normalize the factor data for the fitting
factor_scale <- scale(factor_data) %>% 
  as.matrix()

# fit the normalized factor data to eigenvalues
pca_fit <- factor_scale%*%pca_load %>% 
  as.data.frame()

# calculate the average of PCs to prepare for spatial visualization
pca_fit$GEOID <- full_data$GEOID
pca_fit_tract <- pca_fit %>% 
  group_by(GEOID) %>% 
  summarise(PC1 = mean(PC1, na.rm = TRUE), PC2 = mean(PC2, na.rm = TRUE), PC3 = mean(PC3, na.rm = TRUE), PC4 = mean(PC4, na.rm = TRUE), PC5 = mean(PC5, na.rm = TRUE))

tract <- get_acs(
  geography = "tract",
  variables = "B01001_001",
  state = "PA",
  county = "Allegheny",
  output = "wide",
  geometry = TRUE
)

pca_fit_tract <- left_join(tract, pca_fit_tract, by = c("GEOID" = "GEOID"))
```

```{r}
PA_state_plane <- "+proj=lcc +lat_1=41.95 +lat_2=40.88333333333333 +lat_0=40.16666666666666 +lon_0=-77.75 +x_0=600000 +y_0=0 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"
river <- read_sf("https://openac-alcogis.opendata.arcgis.com/datasets/def3bd39f12d408c9bd097f5a3b3b136_0.geojson?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D") %>% 
  st_transform(PA_state_plane)

ggplot(pca_fit_tract) +
  geom_sf(aes(fill = PC1)) +
   geom_sf(data = river, fill = "yellow", color = NA) 

ggplot(pca_fit_tract) +
  geom_sf(aes(fill = PC2))+
   geom_sf(data = river, fill = "yellow", color = NA) 

ggplot(pca_fit_tract) +
  geom_sf(aes(fill = PC3))+
   geom_sf(data = river, fill = "yellow", color = NA) 

ggplot(pca_fit_tract) +
  geom_sf(aes(fill = PC4))+
   geom_sf(data = river, fill = "yellow", color = NA) 

ggplot(pca_fit_tract) +
  geom_sf(aes(fill = PC5))+
   geom_sf(data = river, fill = "yellow", color = NA) 
```


```{r}
for (i in c(
  "none",
  "varimax",
  "quartimax",
  "equamax",
  "varimin",
  "geominT",
  "Promax",
  "promax",
  "oblimin",
  "simplimax",
  "biquartimin",
  "bifactor",
  "geominQ"
)) {
  filename <- paste0("fa_", i)
    assign(filename,
           fa(correlation,
              nfactors = 4,
              rotate = i))
  }

fit <- data.frame(matrix(rep(NA, 4), nrow=1))[numeric(0), ]
for (i in c(
  "none",
  "varimax",
  "quartimax",
  "equamax",
  "varimin",
  "geominT",
  "Promax",
  "promax",
  "oblimin",
  "simplimax",
  "biquartimin",
  "bifactor",
  "geominQ"
)) {
  for (j in c(1:15)) {
    fa_analysis <- fa(correlation,
                      nfactors = j,
                      rotate = i)
    fa_fit <-
      c(fa_analysis$rotation,
        fa_analysis$factors,
        fa_analysis$fit.off,
        fa_analysis$rms)
    fit <- rbind(fit, fa_fit)
  }
}
colnames(fit) <- c("fa_method", "n_factor","fit_value", "rms")

# rotations do not change the fit value
```


```{r}
fa_oblimin_load <- fa_oblimin$loadings[, 1:4]

fa_oblimin_value <- fa_oblimin$values
fa.diagram(fa_oblimin)
```


```{r}
ggplot(fa_varimax_load) +
  geom_bar(aes(reorder(row.names(fa_varimax_load), MR1), MR1), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# MR1: wealth
```

```{r}
ggplot(fa_oblimin_load) +
  geom_bar(aes(reorder(row.names(fa_oblimin_load), MR2), MR2), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# MR2: commercial area vs residential area
```


```{r}
ggplot(fa_oblimin_load) +
  geom_bar(aes(reorder(row.names(fa_oblimin_load), MR3), MR3), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# CBD
```


```{r}
ggplot(fa_oblimin_load) +
  geom_bar(aes(reorder(row.names(fa_oblimin_load), MR4), MR4), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# diversity, access to home ownership
```

```{r}
# fit the normalized factor data to eigenvalues
fa_fit <- factor_scale%*%fa_oblimin_load %>% 
  as.data.frame()

# prepare for spatial visualization
fa_fit$GEOID <- full_data$GEOID
fa_fit_tract <- fa_fit %>% 
  group_by(GEOID) %>% 
  summarise(MR1 = mean(MR1, na.rm = TRUE), MR2 = mean(MR2, na.rm = TRUE), MR3 = mean(MR3, na.rm = TRUE), MR4 = mean(MR4, na.rm = TRUE))

fa_fit_tract <- left_join(tract, fa_fit_tract, by = c("GEOID" = "GEOID"))
```

```{r}
ggplot(fa_fit_tract) +
  geom_sf(aes(fill = MR1)) +
   geom_sf(data = river, fill = "yellow", color = NA) 

ggplot(fa_fit_tract) +
  geom_sf(aes(fill = MR2))+
   geom_sf(data = river, fill = "yellow", color = NA) 

ggplot(fa_fit_tract) +
  geom_sf(aes(fill = MR3))+
   geom_sf(data = river, fill = "yellow", color = NA) 

ggplot(fa_fit_tract) +
  geom_sf(aes(fill = MR4))+
   geom_sf(data = river, fill = "yellow", color = NA) 
```




